\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles

### 2020-06-29 Response-NAR-Round2.tex

% !TEX TS-program = pdflatex
% !TEX encoding = UTF-8 Unicode

% This is a simple template for a LaTeX document using the "article" class.
% See "book", "report", "letter" for other types of document.

\documentclass[9pt,hyperref]{article} % use larger type; default would be 10pt
\setcounter{secnumdepth}{1}

\usepackage[utf8]{inputenc} % set input encoding (not needed with XeLaTeX)

%\%\% Examples of Article customizations
% These packages are optional, depending whether you want the features they provide.
% See the LaTeX Companion or other references for full information.

%\%\% PAGE DIMENSIONS
\usepackage[margin=1.3in]{geometry} % to change the page dimensions
\geometry{a4paper} % or letterpaper (US) or a5paper or....
% \geometry{margin=2in} \% for example, change the margins to 2 inches all round
% \geometry{landscape} \% set up the page for landscape
%   read geometry.pdf for detailed page layout information

\usepackage{graphicx} % support the \includegraphics command and options

% \usepackage[parfill]{parskip} \% Activate to begin paragraphs with an empty line rather than an indent

%\%\% PACKAGES
\usepackage{hyperref}
\hypersetup{
	colorlinks, linkcolor={red!80!black},
	citecolor={blue}, urlcolor={red!50!black}
}

\usepackage[authoryear]{natbib}
\usepackage{amsmath} % for much better looking tables
\usepackage{xspace}
\usepackage{url}
\usepackage{soul}
\usepackage[table,x11names,dvipsnames]{xcolor}
\usepackage{amssymb} % for much better looking tables
\usepackage{booktabs} % for much better looking tables
\usepackage{array} % for better arrays (eg matrices) in maths
\usepackage{ntheorem} % for better arrays (eg matrices) in maths
\usepackage{paralist} % very flexible & customisable lists (eg. enumerate/itemize, etc.)
\usepackage{verbatim} % adds environment for commenting out blocks of text & for better verbatim
\usepackage{subfig} % make it possible to include more than one captioned figure/table in a single float
\usepackage{mdframed}
\usepackage{relsize}
\usepackage{pgf}
\usepackage{cabin}
\usepackage{enumitem}
\usepackage{tgtermes}
\usepackage{collcell}
\usepackage{adjustbox}% to resize figures

%\usepackage{tgadventor}
% These packages are all incorporated in the memoir class to one degree or another...

%\%\% HEADERS & FOOTERS
\usepackage{fancyhdr} % This should be set AFTER setting up the page geometry
\pagestyle{fancy} % options: empty , plain , fancy
\renewcommand{\headrulewidth}{0pt} % customise the layout...
\lhead{}\chead{}\rhead{}
\lfoot{}\cfoot{\thepage}\rfoot{}

%\%\% SECTION TITLE APPEARANCE
\usepackage{sectsty}
\allsectionsfont{\sffamily\mdseries\upshape} % (See the fntguide.pdf for font help)
% (This matches ConTeXt defaults)

%\%\% ToC (table of contents) APPEARANCE
\usepackage[nottoc,notlof,notlot]{tocbibind} % Put the bibliography in the ToC
\usepackage[titles,subfigure]{tocloft} % Alter the style of the Table of Contents
\renewcommand{\cftsecfont}{\rmfamily\mdseries\upshape}
\renewcommand{\cftsecpagefont}{\rmfamily\mdseries\upshape} % No bold!

\newcommand{\FAnswer}[1]{{\color{black!95}#1}}
\newcommand{\IAnswer}[1]{\FAnswer{\hfill $\to$ #1}}
\newcommand{\PAnswer}[1]{\FAnswer{\\#1}}

\newcommand{\Answer}[1]{\noindent{{{\sffamily Response:} }{\FAnswer{#1}}}}
\newcommand{\Comment}[1]{\item{\color{red!40!black}{{\sffamily Comment:} }{#1}}}
\newcommand{\Quote}[1]{\begin{mdframed}[backgroundcolor=black!2,linecolor=gray!50, linewidth=.6mm]#1\end{mdframed}}

\usepackage[colorinlistoftodos]{todonotes}
\newcommand{\TODO}[2][All]{{\todo[color=blue!10,inline]{\color{black}{TODO (#1)}: #2}}}
\newcommand{\TODOBruno}[2][Bruno]{{\todo[color=red!10,inline]{\color{black}{TODO (#1)}: #2}}}
\newcommand{\TODOAfaf}[2][Afaf]{{\todo[color=Aquamarine3,inline]{\color{black}{TODO (#1)}: #2}}}

%\%\% END Article customizations

%\%\% The "real" document content comes below...

%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%
%\%                                          \%\%
%\% Enter the authors here                   \%\%
%\%                                          \%\%
%\% Specify information, if available,       \%\%
%\% in the form:                             \%\%
%\%   <key>={<id1>,<id2>}                    \%\%
%\%   <key>=                                 \%\%
%\% Comment or delete the keys which are     \%\%
%\% not used. Repeat \author command as much \%\%
%\% as required.                             \%\%
%\%                                          \%\%
%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%\%

\input{macros.tex}

\title{Response to second round of reviews\\[.3em]\OurTool{}:  Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles}
\author{
Afaf Saaidi \and
Delphine Allouche \and
Mireille Regnier \and
Bruno Sargueil \and
Yann Ponty}

\date{} % Activate to display a given date or no date (if empty),
         % otherwise the current date is printed

\begin{document}
\maketitle

\tableofcontents

\section{General comments}

\section{Specific comments}

	\subsection{Reviewer \#2}
	\begin{enumerate}[resume]
\Comment{The revised manuscript is much clearer now. However, I am still puzzled over the performance comparison between IPANEMAP and Rsample over Hajdin et al.’s data. The Hajdin dataset consists of 24 RNAs, but only 16 of them were considered in the Rsample paper (per its supplementary information). As I previously commented, IPANEMAP’s performance gains are really dramatic for the few sequences that were not considered by Spasic et al. For example, 5SrEcoli, signal recognition particle, and E. coli tRNA(phe).

Nevertheless, I also looked at a few of the sequences considered by both Spasic et al. and the authors. Below are a few examples of inconsistencies between these two benchmarks:

[..]

- Can the authors reconcile these discrepancies, or alternatively, am I missing something in the way this benchmark was set up?

Please note the following:
A. I did not check all the RNAs benchmarked by Spasic et al. so there might be additional discrepancies. 
B. Reported values were consistent between the two works for some RNAs which I did review. 
C. Given these discrepancies, I am not sure what to conclude from the performances reported for those sequences benchmarked in this manuscript but not by Spasic et al.

Maybe the predictions of these RNAs are very sensitive to the thermodynamic model parameters, I don’t know. If this is indeed the case, it is a good idea to benchmark IPANEMAP against more than a single method, similarly to what Spasic et al. did. I think this creates a more balanced picture of current structure prediction capabilities.}

    \Answer{We are greatly indebted to this reviewer for his/her}

	\Comment{Additionally, referring to the authors comment on page 24: “we object to such an adversarial view of this particular analysis (and, more generally, our work)…”  I would like to clarify I had no such intention. My critique aimed at strengthening or solidifying the reported results. I am sorry the authors interpreted my requests for better controls and comparisons as a negative opinion of their work.}

	\Answer{Actually, in addition to being wrong regarding the benchmark, we expressed ourselves poorly in this instance: Having done more than our fair share of reviewing and editorial activities...}
	\end{enumerate}

\section{Referee: 3}

	\subsection{Minor Revisions:}
		\begin{enumerate}[resume]
	\Comment{On page 4, in “Cordero dataset”, the text states that the reactivity is set to 0 for nucleotide identities that are unreactive to a specific probing agent.  But 0 is a reactivity that receives a negative pseudoenergy free change, encouraging base pairing.  It would be much better to give a 0 pseudo free energy change.  Is it possible the manuscript should really state a 0 free energy change?  If not, the authors should add some text to justify this choice.}

    \Answer{WE set the pseudo-energy to 0 (and not the reactivity, as erroneously stated in the manuscript). \TODO{correct}}

	\Comment{On page 4, the text states “a standard denaturation protocol”, but I think the manuscript should state “a standard annealing protocol”.  It looks like the intent was to anneal the structure to find its equilibrium structure at 37 degrees C, not to leave the structure denatured.}

    \Answer{Indeed, real denomination is denaturation/renaturation protocol, currently the standard nomenclature. \TODO{correct}}
	\Comment{On page 5, in statistical significance, “usual significance threshold” should really state the “usual type I error rate”.  Alpha is the probability of falsely rejecting the null hypothesis (type I error).  “Threshold” doesn’t convey this meaning.  }

    \Answer{\TODO{correct}}
	\Comment{On page 8, “barely significant” should just be changed to “but significant”. The improvements may be modest, but they are significant.  I don’t agree that something can be “barely” significant.}

    \Answer{\TODO{be polite}}
	\Comment{On page 10, “Considering three conditions yields improvement over the average performance of the triplet” confuses me.  I think what was intended is “Considering three conditions yields improvement over the average performance of the doublet [/pair]” }
    \Answer{\TODO{The average performance of prediction informed by individual conditions is lower than that of predictions simultaneously informed by the three conditions}}

	\subsection{Typos:}
	\Comment{On page 1, “different dynamics, that can help” should be “different dynamics that can help”.}

	\Answer{}
	\Comment{On page 3, “using an iterative heuristics” should be “using an iterative heuristic”.  Also, “we elect the most promising cluster” should be “we select the most promising cluster”.}

	\Answer{}
	\Comment{On page 6, “Spacic” should be “Spasic”.}

	\Answer{}

	Suggestions:
	\Comment{On page 1, there is a new phrase about probes “some of which are usable in vivo”.  My recollections is that Zaug and Cech were the first to demonstrate in vivo probing of RNA structure, and it is probably worth citing the paper [RNA. 1995 1:363.].}

	\Answer{\TODO{Ok Paillard PMID 15355993}}
	\Comment{Page 1 claims that using multiple data sources has never been fully automated.  I agree that it has never been automated when using quantitative data and soft constraints.  On the other hand, the deprecated hard constraint method (reference 2) was fully automated and used multiple probing methods.  The phrase should be made more specific.}

	\Answer{\TODO{}}

	\end{enumerate}

\bibliographystyle{abbrvnat}
\bibliography{biblio}
\end{document}

### 2020-06-29 IPANEMAP-2019.tex

\paragraph{Hajdin dataset.} 
A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions. It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) is available.
This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures.

\paragraph{Cordero dataset.} 
In order to test the integration of different probing sources, we consider a dataset introduced in Cordero\etal\cite{Cordero2012}, consisting of $6$ RNAs with known structures, for which reactivities are available for three reagents: {\tt NMIA-\SH, DMS} and {\tt CMCT}. 
Probing data were downloaded from the RMDB~\citep{Cordero2012a} on July 2017. In the RMDB, reactivity scores are reported for all nucleotides, including those that are not expected to react with a given reagent. Thus, for the DMS (resp. CMCT) probing, we restricted reactivities to positions featuring nucleotides {\sf A} and {\sf C} (resp. {\sf G} and {\sf U}), setting the reactivity to 0 for other positions. This assumption allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).

@@

The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, based on its agreement with an accepted reference (native) structure $\Ref$. It can be interpreted as a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as 
$$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
where TP, FP, TN, FN classically represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.

@@

\Rev{
\subsection*{Assessment of statistical significance}
Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances, using a two-tailed Student $t$-test with unequal variance, using a classic significance threshold $\alpha=5\%$. When comparing performances within a given dataset, we used the paired version of the test.

Intuitively, the $t$-test estimates the probability that an observed change in performance can be attributed to chance alone.
It considers the hypothesis $H_0$ that the two distributions have equal mean and computes a \Def{P-value}, \emph{i.e.} the likelihood of the observed values under $H_0$. If the P-value is lower than $\alpha$, then $H_0$ is deemed \Def{refuted}, and the observed difference in average performance is considered as significant.

@@

Since \OurTool{} supports any number of probing profiles, we considered the Cordero\etal\cite{Cordero2012a} dataset, further described in the previous section. It consists of 6 RNAs of known structures, for which reactivity profiles are available for the CMCT, NMIA and DMS reagents. We executed \OurTool{} with default parameters on each sequence and any subset of the three available conditions. The centroid secondary structure  associated with the largest probability cluster was considered as the final prediction. Predictions were evaluated in term of the geometric mean (GM) metric, for the sake of comparison with previous studies~\cite{Spasic2017}. As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes. Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data (Detailed results can be found in Supp. Tables S1 and S2).

@@

\Rev{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).  For all RNAs, a single structure was returned by both software. Note that the Pareto-optimality always implies a single returned cluster/structure for \OurTool{} in a mono probing setting.}
We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) associated with predicted structures. Sensitivity and PPV reported in Supp. Table S4, while Supp. Table S5 reports the stability of predictions across 10 runs of \OurTool.

@@

This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling. However, the two methods differ substantially whenever a single conformation is detected, in which case \Software{RSample} switches to a simple MEA prediction, whereas \OurTool{} always proceeds with an iterative clustering. This allows \OurTool{} to consider a larger number of clusters, which we believe helps eliminating structural outliers, leading to the election of a better final centroid structure.

@@

As described in the Material and Methods section, we probed \didy{}  in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites. These experimental results were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.

\paragraph{Comparison and clustering of conditions.}
Faced with diverse probing conditions, we first assessed the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control. We used the methodology described in the Material and Methods~\ref{sec:dotplots} and, for each condition, computed the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble induced by the reactivity profile. We then computed the Ensemble Distance, the squared Euclidean distance between dot plots, for each pair of conditions. Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis.

@@

Across the sixteen conditions, we observed a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE). \Rev{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp. Table S6).}
 These predictive capacities clearly outperformed those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of predictions of good overall quality. Indeed, it is worth reminding that the MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor. The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.

@@

\paragraph{Bi-probing mitigates the influence of outliers, but does not significantly improve prediction quality.}
Next, we turned to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempted to quantify their impact on \OurTool predictions. For each pair and triplets of conditions, we executed \OurTool using 1~000 structures per condition, and considered the first returned structure. We then computed the associated MCC, and compared it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments. A summary of the results over pairs is shown in Figure~\ref{fig:pairwise} (see details in Supp. Table S7).

@@

To test this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Table~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp. Figure~S1 for associated models). 
\Rev{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}

@@

	\caption{Predictive performances of conditions across clusters.  In each line, a single condition is chosen from each cluster except for cluster \Bull{B}. \Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average MCC of mono conditions, in all cases in presence/absence of \NAICE.}

@@

\subsection*{Supplementing wild-type profiles with random single mutants increases prediction quality}
\Rev{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
The joint analysis of several mutants may appear ill-advised, as single point mutants are expected to adopt different structures than the wild-type sequence (WT) , an expectation that is at the core of downstream analyses~\cite{Cordero2015}. However, those changes are typically local, and the limited influence of outliers on \OurTool{} substantiates the hope that the benefits of including independently-produced conditions will outweigh the cost of noise, introduced by local changes.}

We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we each supplemented with the WT sequence.
For each set of probing profiles (\Cond{DMS}{mg}{MaP}{ilu}{}), we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metric. By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions. We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool (see Supp. Figure S2 for details).

@@

\Rev{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\cite{Janssen2015}, or a iterative heuristic alternative~\cite{Hajdin2013}. The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis. However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble. A similar behavior is observed for the other datasets considered in our study (see Supp. Tables S2 and S5). %In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
Finally, the current clustering method induce a growth of time and memory requirements that scales quadratically with the number of samples. While this allows a joint analysis of a dozen of reactivity profiles in less than an hour on a personal computer, we expect the ability to push the number of samples will lead to better, even more reproducible, predictions. To that purpose, we will explore embedding techniques and linear-time community detection algorithms as alternatives to MBkM~\cite{Sculley2010}.}

>>>>

\paragraph{Hajdin dataset.} 
A dataset was gathered by Hajdin\etal\cite{Hajdin2013} to validate the predictive capacities of probing data-driven predictions. It consists of $24$ RNA sequences with known secondary structures for which a single chemical probing reactivity profile (1M7-\SH) was made available.
This dataset includes sequences originating from a variety of organisms, and spans lengths ranging from $34$ nts to $530$ nts, with a focus on riboswitches and complex RNA architectures \Rev{(full list in Supp. Table S3)}.

\paragraph{Cordero dataset.} 
In order to test the integration of different probing sources, we consider a dataset introduced in Cordero\etal\cite{Cordero2012}, consisting of $6$ RNAs with known structures, for which reactivities are available for three reagents: {\tt NMIA-\SH, DMS} and {\tt CMCT}. 
Probing data were downloaded from the RMDB~\citep{Cordero2012a} on July 2017. In the RMDB, reactivity scores are reported for all nucleotides, including those that are not expected to react with a given reagent. Thus, for the DMS (resp. CMCT) probing, we restricted reactivities to positions featuring nucleotides {\sf A} and {\sf C} (resp. {\sf G} and {\sf U}), setting the reactivity to 0 for other positions. This allowed to decrease the noise generated by reactivities associated with non-targeted nucleotides, leading to more accurate predictions (data not shown).

@@

The Matthews Correlation Coefficient (MCC) is a classic metric for assessing the quality of a predicted structure $S$, identified by a set of base pairing positions, compared to an accepted reference (native) structure $\Ref$. It represents a compromise \Rev{between the main metrics derived from the confusion matrix}, and is defined as 
$$\text{MCC}(S\mid \Ref)=\frac{\text{TP} \times \text{TN} -\text{FP} \times \text{FN}}{\sqrt[]{(\text{TP}+\text{FP})(\text{TP}+\text{FN})(\text{TN}+\text{FP})(\text{TN}+\text{FN})}},$$
where TP, FP, TN, FN represent the correctly/erroneously predicted and correctly/erroneously omitted base pairs in $S$ with respect to $\Ref$.
\Rev{MCC is a stringent metric, taking values between -100\% and 100\%, 0\% being the expected MCC of a "coin tossing" random predictor.}

@@

\Rev{
\subsection*{Assessment of statistical significance}
Following Xu\etal\cite{Xu2012}, we assessed the statistical significance of observed changes in predictive performances using a two-tailed Student $t$-test (unequal variance) with usual significance threshold $\alpha=5\%$. When comparing performances within a given dataset, we used the paired version of the test.

Intuitively, the $t$-test estimates the probability that an observed change in performance can be attributed to chance alone.
It considers the hypothesis $H_0$ that the two distributions have equal mean and computes a \Def{P-value}, \emph{i.e.} the likelihood of the observed data under $H_0$. If the P-value is lower than $\alpha$, then $H_0$ is deemed \Def{refuted}, and the observed difference in performances is considered significant.

@@

Since \OurTool{} supports any number of probing profiles, we considered the Cordero\etal\cite{Cordero2012a} dataset, further described in the previous section. It consists of 6 RNAs of known structures, for which reactivity profiles are available for the CMCT, NMIA and DMS reagents. We executed \OurTool{} with default parameters on each sequence and any subset of the three available conditions. The centroid secondary structure  associated with the largest probability cluster was considered as the final prediction. Predictions were evaluated in term of the geometric mean (GM) metric, for the sake of comparison with previous studies~\cite{Spasic2017}. As a control experiment, we also report ${\tt RNAfold}$ predictions in presence/absence of probing data, both in energy minimization (MFE) and maximum expected accuracy (MEA) modes. Figure~\ref{cordero1} shows the averaged GM values for all combinations of tools and probing data (Detailed results in Supp. Tables S1 and S2).

@@

\Rev{We used the default number of sampled structures for \Software{RSample}, namely 10~000 to compute correction factors and 1~000 for the clustering phase, and \OurTool (1~000 structures).  For all RNAs, a single structure was returned by both software. Note that the Pareto-optimality always implies a single returned cluster/structure for \OurTool{} in a mono probing setting.}
We computed and report in Figure~\ref{fig:Vsrsample} the individual Geometric Means (GM) of predicted structures. \Rev{Sensitivity and PPV can be found in Supp. Table S4, and Supp. Table S5 reports the stability of predictions over 10 independent runs of \OurTool.}

@@

This observed discrepancy is somewhat surprising, given that \OurTool and \Software{RSample} share the same sampling/clustering based modeling. However, the two methods differ substantially whenever a single conformation is detected, in which case \Software{RSample} switches to a simple MEA prediction, whereas \OurTool{} always proceeds with an iterative clustering. This allows \OurTool{} to consider a larger number of clusters, which we believe helps eliminating structural outliers, leading to the election of an improved final centroid structure.

@@

As described in the Material and Methods section, we probed \didy{}  in 16 different conditions using different SHAPE and classical reagents, in presence/absence of Mg$^\text{2+}$ ions, using either the standard premature stop or the newly developed mutation induction technologies to map the modification sites. \Rev{We observed an overall good agreement of reactivities with the 3D model proposed by Meyer~\etal\cite{Meyer2014} (see  Supp. Figures S1, S2 and S3).} Reactivities were then used as input of \OurTool{}, either individually or combined,  for secondary structure prediction.

\paragraph{Comparison and clustering of conditions.}
Faced with diverse probing conditions, we first assessed the compatibility of the conclusions drawn from different probing data, including the probing-free run of \OurTool as a control. We used the methodology described in the Material and Methods~\ref{sec:dotplots} and, for each condition, computed the base pair probability distribution (aka \emph{dot-plot}) in the  pseudo-Boltzmann ensemble induced by the reactivity profile. We then computed the Ensemble Distance, the squared Euclidean distance between dot plots, for each pair of conditions. Figure~\ref{fig:PCA} summarizes the pairwise distance, projected onto a 2D surface by a principal component analysis. \Rev{A systematic cross-analysis of raw reactivities revealed a positive, if modest, correlation between conditions, but did not support a clear partition of profiles based on probing technology or presence/absence of Mg$^{\text{2+}}$ (see Supp. Figure S4).}

@@

Across the sixteen conditions, we observed a large discrepancy in the capacity of the experimental setup to inform predictions, with MCCs ranging from above 80\% (\NMIA, \NMIAMg, \OneMSevILU and \OneMSevILUMg) to around 60\% (\NAICE, \BzCN, \OneMSevCE and \NMIACE). \Rev{Interestingly the MCC value cannot be trivially anticipated from the compatibility of the reactivity scores with the predicted structure, or even with the native structure (see Supp. Table S1).}
 These predictive capacities clearly outperformed those achieved in the absence of probing data (51\% MCC), and can be interpreted as indicative of predictions of good overall quality. The lowest MCC value of 60\% is, in particular, equally consistent with Sensitivity/PPV values of 60\%/60\%, 80\%/46\% or 45\%/82\%.

@@

\paragraph{Bi-probing mitigates the influence of outliers, but does not significantly improve prediction quality.}
Next, we turned to a systematic exploration of the predictive capacities of bi-probing analyses, based on pairs of probing profiles, and attempted to quantify their impact on \OurTool predictions. For each pair and triplets of conditions, we executed \OurTool using 1~000 structures per condition, and considered the first returned structure. We then computed the associated MCC, and compared it with the MCC of the worst performing condition (Min), best-performing condition (Max) and average MCC over the single conditions experiments. A summary of the results over pairs is reported in Figure~\ref{fig:pairwise} (details in Supp. Table S6).

@@

To test this hypothesis, we executed \OurTool{} on every subset of conditions within clusters, and report in Table~\ref{fig:similarconds} the quality (MCC) of predictions for any subset of similar conditions (see Supp. Figure~S5 for associated models). 
\Rev{Note that such an analysis does not use knowledge of the native structure, since our clustering only relies on properties of the pseudo-Boltzmann ensemble.}

@@

	\caption{Predictive performances of conditions across clusters.  A single condition is chosen from each cluster except for \Bull{B}. \Rev{Clusters consisting of a single condition (\Bull{D}, \Bull{F} and \Bull{H}) contribute additional conditions  $ :=\{\OneMSevILUThree, \BzCNMg, \DMSMg\}$.
		The MCC of the predicted structure is reported ('Joint' columns) and compared to the average of mono conditions ('Mono' columns), in presence/absence of \NAICE.}

@@

\subsection*{Supplementing wild-type profiles with random single mutants increases prediction quality}
\Rev{In this final analysis, we tested the capacity of \OurTool to extract and exploit complementary information produced by the systematic probing of single-point mutants using the Mutate-and-Map (MaM) protocol~\cite{Cordero2015}.
The joint analysis of several mutants may appear error-prone, as single point mutants are expected to adopt different structures than the wild-type sequence (WT), an expectation that is at the core of downstream analyses~\cite{Cordero2015}. However, such changes are typically local, and the robustness of \OurTool{} to outlier conditions nourishes the hope that the benefits of including independently-produced conditions will outweigh the cost of noise introduced by local changes.}

We generated 100 uniformly-distributed subsets consisting of 1, 2, 3 and 10 mutants extracted from the Cheng dataset~\cite{Cheng2017}, which we each supplemented with the WT sequence.
For each set of probing profiles (\Cond{DMS}{mg}{MaP}{ilu}{}), we executed \OurTool{} with default parameters, using a sample size of 1~000 structures. 
We also analyzed the WT sequence alone, reproducing the analysis 100 times to compare the variability across sets of mutants to the one induced by stochastic aspects of our method. 
Finally, we performed a restricted analysis focusing on the WT, supplemented with 2 probing profiles selected from the 20 most similar conditions to the WT , as assessed by the ensemble distance metric. By comparing our results with those obtained for unrestricted pairs of mutants, we assess the influence of structure variability on the quality of our predictions. We report in Figure~\ref{fig:variantanalysis} the distribution of MCC values associated with the first prediction returned by \OurTool (see Supp. Figure S6 for details).

@@

\Rev{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\cite{Janssen2015}, or a iterative heuristic alternative~\cite{Hajdin2013}. The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis. However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as \emph{focusing} the predictions onto a subset of the Boltzmann ensemble. A similar behavior is observed for the other datasets considered in our study (see Supp. Tables S2 and S5). %In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
Finally, the current clustering method induce a growth of time and memory requirements that scales quadratically with the number of samples. While this allows a joint analysis of a dozen of reactivity profiles in less than an hour on a personal computer, we expect the ability to push the number of samples will lead to better, even more reproducible, predictions. To that purpose, we will explore embedding techniques and linear-time community detection algorithms as alternatives to MBkM~\cite{Sculley2010}.}

Future extensions also include \Rev{addressing} the dilemma of whether or not to include Mg$^{\text{2+}}$ ions during the probing, in the perspective of a downstream computational prediction. On the one hand, RNA reaches its native 3D structure in the presence of Mg$^{\text{2+}}$ ions~\cite{Knapp1989}. However, such native structures include non WC pairings, and complex topological motifs such as pseudoknots, kissing loops, tetraloop interactions, that classic structure prediction algorithms do not explicitly capture. In contrast, in the presence of only monovalent ions, RNA is expected to form its helical domains, but leave its  tertiary contacts unstable, with the notable exception of the G quadruplex. Therefore, probing the RNA without Mg$^{\text{2+}}$ could provide a signal that is less deceptive, and thus more informative, to predicting algorithms. Unfortunately, destabilizing tertiary contacts increase the probability of misfolding, or the coexistence in solution of multiple conformations. In this work we probed \didy, which contains many tertiary motifs, in the presence and absence of Mg$^{\text{2+}}$. Remarkably, the most accurate predictions were obtained with 1M7 and NMIA \Rev{irrespectively of the presence/absence of Mg$^{\text{2+}}$, and despite the fact that reactivity profiles substantially diverged between the two conditions (see Supp. Figure S4)}. However, for other reagents/protocols, predictions appeared positively impacted (+$10\%$ average MCC) by the presence of Mg$^{\text{2+}}$, painting a mixed picture that would deserve future investigations.
%\item je pense qu'il serait intéressant au moins pour un uo deux réactif de voir ce que donne le couple + et - Mg. Peut etre par exemple avec le 1M7ilu et un autre mauvais genre le NAI
%\end{itemize}
% il faudrait regrder ce que fait weeks avec le 1M7, le NMIA et le 1M6, je n'ai pas bien compris quelle tambouille il fait, mais ca part du meme genre de constatation : il y a une spécificité dans chacune des sondes qu'il n'arrive pas vraiment à formaliser donc il fait une tambouille qui lui permet de tous les intégrer (mais c'est peut etre un peu pédestre comme démarche!). Apres ce sont probablement des conditions tres proches ce qui n'apporte pas d'apres nos résultats bcq d'infos supplémentaires .. bref ca vaudrait peut etre le coup de comparer.
